Planning to Save Energy: How Information Format Affects Accuracy

### 2024-12-08 manuscript/manuscript.qmd
big update to results section. Added a few paragraphs to Intro.

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
format:
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    html: 
      echo: true
      code-fold: show
      output-file: manuscript-code.html
    hugo-md:
        include: false
        echo: false
        html-math-method: mathjax
        output-file: manuscript_hugo.md
    gfm:
        echo: true
        output-file: manuscript_gfm.md
    hikmah-manuscript-pdf:
        output-file: rep_study.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    hikmah-manuscript-docx: default
format-links:   
    - text: HTML
      href: manuscript.html
      icon: file-minus
    - text: HTML (with code)
      href: manuscript-code.html
      icon: file-plus
    - format: hikmah-manuscript-pdf
      text: PDF version
      icon: file-pdf
    - hikmah-manuscript-docx
    - gfm
date: 2024-12-07
abstract: |
    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans. By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
---

@@

- @canfieldPerceptionsElectricityuseCommunications2017 found that tables were more effective than graphs for conveying specific electricity usage data, likely because tables facilitate straightforward point reading. However, they also noted that individuals with lower energy literacy had reduced comprehension across all formats.
- @canfieldPerceptionsElectricityuseCommunications2017's findings on preferences for historical use information and the impact of neighbor comparisons
- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements [@vesseyCognitiveFitTheoryBased1991]
- alignment can reduce cognitive load and enhance accuracy in planning [@shahBarLineGraph2011]
- Reimer et al. (2015) provide  context on how numerical formats affect risk perception, the reference class problem, and the benefits of natural frequencies.

@@

@@

We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk. In Experiment 1, 252 participants were intially recruited, but data from 17 participants were corrupted due to experimenter error, leaving a final sample of 235 participants. Most participants (76%) reported using a calculator to complete the task.

@@

Participants received energy usage data for two hypothetical families and were tasked with creating action plans to meet specified reduction goals by allocating usage across five appliance categories:  heating, cooling, water heating, refrigerator, and an other appliances (e.g., TV, lighting).

For each family scenario, the particpipants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh). For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task). Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S. dollars. In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.

@@

```{r}
pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

@@

```{r}
#| label: tbl-s1-agg
#| tbl-cap: "Study 1: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# proportion matching vs close match

# proportion matching vs close match

# overall pct of subjects who matched their goal
s1_agg4 |> group_by(refClass) |>
    summarise(
    'Avg. % meeting goal' = mean(pct),
    'Avg. Deviation From Goal' = median(mean_pct_change),
    # sd = sd(pct),
    # n = n(),
    #se=sd(pct)/sqrt(n)
) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

@@

For our primary analyses of participants' ability to create accurate energy-saving plans, we employed an accuracy level binning approach by categorizing responses into four distinct levels: Exact match, 0.01–2% error, 2.01–15% error, and Over 15% error. The current analysis employs a cumulative ordinal regression model, implemented via a Bayesian hierarchical framework [@burknerBrmsPackageBayesian2017] . This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category. All analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the tidyverse package [@wickhamWelcomeTidyverse2019].

@@

see @fig-s1-plot

@@

We analyzed planning accuracy using Bayesian ordinal regression. The dependent variable, plan error, was computed by binning the goal deviation into four ordered levels: exact match (0% error), minor deviations (0.01-2% error), moderate deviations (2.01-15% error), and major deviations (>15% error). For each comparison, we provide posterior odds ratios (OR) and their 95% CIs.

```{r}
#| label: tbl-s1-ord
##| tbl-cap: "Study 1: Ordinal Regression Model Results."

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass + (1|id) + (1|state),
    data = s1_agg,
    family = cumulative("logit"),
    cores = 4,
    iter = 5000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
)

#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

@@

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, Category=.category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

#plot(conditional_effects(ordinal_model_s1),points=TRUE)

# or_table |>
#       rownames_to_column(var = "Term") |>
#       filter(!stringr::str_detect(Term, "Intercept")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))

# or_table <- data.frame(exp(fixef(ordinal_model_s1)))
# or_table |>
#   rownames_to_column(var = "Term") |>
#   filter(!stringr::str_detect(Term, "Intercept")) |>
#   as_flextable() |>
#   set_header_labels(values = list(
#     OR = "Odds Ratio",
#     `2.5 %` = "Lower 95% CI",
#     `97.5 %` = "Upper 95% CI"
#   )) |>
#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
#   align(align = 'center', part = 'all') |>
#   fontsize(size = 10, part = "all")
```

::: {#tbl-s1-or}
|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
|:------------------|:--------|:--------|:--------|:----|
|Intercept[1]       |-1.94    |-3.19    |-0.74    |1.00 |
|Intercept[2]       |0.20     |-1.02    |1.42     |0.63 |
|Intercept[3]       |4.36     |3.11     |5.65     |1.00 |
|refClassPercentage |0.85     |-0.54    |2.23     |0.88 |
|refClassUSD        |2.72     |1.41     |4.06     |1.00 |

**Experiment 1**: Ordinal Regression results.

@@

::: {#tbl-s1-epred}
|Term               |Estimate |Est.Error |Q2.5 |Q97.5 |
|:------------------|:--------|:---------|:----|:-----|
|refClassPercentage |2.3      |2         |0.58 |9.3   |
|refClassUSD        |15.2     |2         |4.11 |58.1  |

**Experiment 1**: Odds ratios for group comparisons.

@@

The ordinal model is parameterized with thresholds (intercepts), and positive coefficients can indicate that it is more difficult to achieve higher accuracy categories in the USD condition. 
The model output suggests that, compared to the kWh condition, the USD condition shows a positive coefficient (Estimate = 2.72, 95% CI: 1.41 to 4.06) for the ordinal outcome. At least to me, this positive coefficient appears to indicate that, relative to the kWh reference class, participants in the USD condition are more likely to fall into higher numerical categories of the dependent variable coding. However, because the dependent variable is ordered from best (Exact match) to worst (Over 15% error), care is needed in interpretation. The Percentage condition coefficient (Estimate = 0.85, 95% CI: -0.54 to 2.23) is more uncertain, with its credible interval overlapping zero.  Posterior predictive checks (Figure 3) showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-plot2).

```{r}
#| label: fig-s1-plot2
#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal.  A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
#| fig.width: 11
#| fig.height: 7

#bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])

# Plot predicted probabilities
# ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = Category)) +
#     geom_bar(stat = "identity", position = position_dodge()) +
#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
#                     position = position_dodge(width = 0.9), width = 0.2) +
#     labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
#     ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
#     theme_minimal()

#pp_check(ordinal_model_s1) + ggtitle("Study 1: Posterior Predictive Checks for Ordinal Regression Model") + theme_minimal()

@@

@@

```{r}
#| label: tbl-s2-agg
#| tbl-cap: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.001)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

>>>>

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
format:
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    html: 
      echo: true
      code-fold: show
      output-file: manuscript-code.html
    hugo-md:
        include: false
        echo: false
        html-math-method: mathjax
        output-file: manuscript_hugo.md
        suppress-bibliography: true
    gfm:
        echo: true
        output-file: manuscript_gfm.md
        suppress-bibliography: true
    hikmah-manuscript-pdf:
        output-file: rep_study.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    #hikmah-manuscript-docx: default
    docx: default
format-links:   
    - text: HTML
      href: manuscript.html
      icon: file-minus
    - text: HTML (with code)
      href: manuscript-code.html
      icon: file-plus
    - format: hikmah-manuscript-pdf
      text: PDF version
      icon: file-pdf
    - docx
    - gfm
date: 2024-12-07
abstract: |
    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans. By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
---

@@

@@

The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements [@vesseyCognitiveFitTheoryBased1991], and that such an alignment can reduce cognitive load and enhance accuracy in planning [@shahBarLineGraph2011]. For instance, tables are generally more effective than graphs for conveying specific electricity usage data because they facilitate straightforward point reading [@canfieldPerceptionsElectricityuseCommunications2017]. However, the effectiveness of the format varies with the type of information and individual differences, such as energy literacy, which significantly impacts comprehension and conservation intent. Moreover, the unit in which numerical information is presented influences how decision-makers evaluate and choose between options, with default units increasing value sensitivity [@herberzKiloWhatDefault2020]. In the context of energy, presenting information in terms of multiple translations can increase preference for options aligned with activated objectives, such as pro-environmental values [@ungemachTranslatedAttributesChoice2018]. Furthermore, mental accounting mechanisms, where individuals create mental budgets linking specific consumption acts to specific payments, significantly impact energy decisions and behaviors [@hahnelMentalAccountingMechanisms2020].

Evidence from research on energy consumption feedback, normative comparisons, and eco-feedback platforms suggests that comprehensible and contextually meaningful data presentations can improve users’ ability to plan reductions, especially when these formats are integrated into daily routines [@canfieldPerceptionsElectricityuseCommunications2017; @fischerFeedbackHouseholdElectricity2008; @kimMySmartEEcofeedbackGaming2022; @schwartzWhatPeopleConsumption2015]. Furthermore, temporal and monetary frames have been shown to alter decision quality, with monthly costs or absolute consumption levels often encouraging more energy-efficient intentions than abstract annual or percentage-based metrics [@gillConsumerMatchingCosts2022; @larrickMPGIllusion2008]. In this context, tailoring reference classes to align with intuitive cognitive processes can help bridge the gap between aggregate reduction goals and targeted, appliance-specific conservation strategies.

@@

We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk. In Experiment 1, 252 participants were initially recruited, but data from 17 participants were corrupted due to experimenter error, leaving a final sample of 235 participants. Most participants (76%) reported using a calculator to complete the task.

@@

Participants received energy usage data for two hypothetical families and were tasked with creating action plans to meet specified reduction goals by allocating usage across five appliance categories:  heating, cooling, water heating, refrigerator, and and other appliances (e.g., TV, lighting).

For each family scenario, the participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh). For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task). Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S. dollars. In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.

@@

### Data Analysis

All preprocessing and analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the tidyverse package [@wickhamWelcomeTidyverse2019]. Mixed Bayesian regressions were fit using the brms package [@burknerBrmsPackageBayesian2017], with participants and family scenario (states) set as random effects.

```{r}
#| cache: false
pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
    ggplot2,gt,knitr,kableExtra,ggh4x,patchwork, ggridges,ggstance,lme4,flextable,pander)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))

theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

@@

```{r}

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

```{r}
#| eval: false
##| label: tbl-s1-agg
##| tbl-cap: "Study 1: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

# overall pct of subjects who matched their goal
s1_agg4 |> group_by('Reference Class' = refClass) |>
    summarise(
    'Avg. % Change' = mean(mean_pct_change),
    '% meeting goal (exact)' = mean(pct),
    '% meeting goal (close match)' = mean(pct_close),
    'Abs. Deviation' = median(mean_abs_error),
    'Log Abs. Deviation' = (median(mean_log_abs_error)),
    # sd = sd(pct),
    # n = n(),
    #se=sd(pct)/sqrt(n)
) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

@@

::: {#tbl-s1-agg}

|Reference Class |Avg. % Change |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
|:---------------|:-------------|:----------------------|:----------------------------|:--------------|:------------------|
|kWh             |0.22          |0.38                   |0.54                         |0.03           |-3.7               |
|Percentage      |0.21          |0.22                   |0.40                         |0.06           |-3.1               |
|USD             |0.23          |0.10                   |0.22                         |0.10           |-2.4               |

Study 1: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal

@tbl-s1-agg that participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition. Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than percentages or USD.

As shown in @tbl-s1-agg, participants in the kWh condition exactly met the target reduction goal 38% of the time, significantly outperforming those in the Percentage (22%) and USD (10%) conditions. Furthermore, the kWh reference class exhibited notably smaller mean absolute deviations (0.03) compared to Percentage (0.06) and USD (0.10), suggesting that presenting the reduction goal in absolute units facilitated more precise allocations.

```{r}
#| label: fig-s1-log-dist
#| fig-cap: "Study 1: Distribution of log absolute error by reference class."
#| fig-width: 11
#| fig-height: 8

s1_ld <- ggplot(s1_agg, aes(y = refClass, x = log_abs_error, color = refClass)) +
  geom_density_ridges(aes(fill = refClass), alpha = 0.2, scale = 0.5,
                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
                      position = position_raincloud(width = 0.05, height = 0.1,
                                                    ygap = 0.05)) +
  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
 # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
  guides(fill = "none", color = "none") +
  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "Reference Class") +
  theme(legend.position = "top")

s1_ldc <- ggplot(s1_agg, aes(y = refClass, x = log_abs_error, color = calc)) +
  geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
                      jittered_points = TRUE, point_alpha = 0.7, point_size = .4,
                      position = position_raincloud(width = 0.05, height = 0.1,
                                                    ygap = 0.05)) +
  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  guides(fill = "none") +
  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "") +
  theme(legend.position = "top")

s1_ld | s1_ldc

@@

@@

We next categorized responses into four accuracy levels (exact match [0% error], minor deviations [0.01–2%], moderate deviations [2.01–15%], and major deviations [>15%]) for our primary statistical modeling. Using Bayesian ordinal regression, we modeled the ordered accuracy outcome as a function of the reference class condition, while controlling for random variation across participants and family scenarios:

$$
\text{Accuracy Level} \sim \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})
$$

This approach allowed us to estimate thresholds (intercepts) and regression coefficients that capture how different reference classes affect the likelihood of achieving higher accuracy categories. For each comparison, we provide posterior odds ratios (OR) and their 95% CIs. This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.

```{r}
#| eval: true
##| label: tbl-s1-ord
##| tbl-cap: "Study 1: Ordinal Regression Model Results."

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass +calc + (1|id) + (1|state),
    data = s1_agg_filtered,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_grpclean_calc_2.5.rds')) 
)

@@

# Get predicted probabilities
# pred_summary <- ordinal_model_s1 |>
#     epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
#                 ndraws = 1000, re_formula = NA) |>
#     group_by(refClass, Category=.category) |>
#     summarise(
#         mean_prob = mean(.epred),
#         lower_ci = quantile(.epred, 0.025),
#         upper_ci = quantile(.epred, 0.975)
#     )
#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

# odds ratios of fixed effects
# as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
#     rownames_to_column(var = "Parameter") %>%
#     mutate(across(where(is.numeric), exp)) |>
#     filter(!stringr::str_detect(Parameter, "Intercept")) |> 
#     filter(!stringr::str_detect(Parameter, "calc")) |> 
#     # rename columns to |comparison           | odds_ratio| ci_lower| ci_upper|
#     rename(Comparison = Parameter, odds_ratio = Estimate, ci_lower = Q2.5, ci_upper = Q97.5) |>
#     # rename levels to Percentage vs kWh; USD vs kWh; 
#     mutate(Comparison = case_when(
#         str_detect(Comparison, "refClassPercentage") ~ "Percentage vs kWh",
#         str_detect(Comparison, "refClassUSD") ~ "USD vs kWh",
#         TRUE ~ Comparison
#     )) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))

::: {#tbl-s1-reg}

|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
|:------------------|:--------|:--------|:--------|:----|
|Intercept[1]       |-3.8     |-5.45    |-2.28    |1.00 |
|Intercept[2]       |-1.7     |-3.29    |-0.15    |0.98 |
|Intercept[3]       |2.8      |1.27     |4.40     |1.00 |
|refClassPercentage |1.3      |0.01     |2.66     |0.98 |
|refClassUSD        |2.8      |1.52     |4.04     |1.00 |
|calcUsedCalculator |-2.8     |-4.09    |-1.56    |1.00 |

**Experiment 1**: Ordinal Regression results. Ordinal regression results. Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline.

@@

::: {#tbl-s1-ord}
|Comparison        |odds_ratio |ci_lower |ci_upper |
|:-----------------|:----------|:--------|:--------|
|Percentage vs kWh |3.7        |1.0      |14       |
|USD vs kWh        |15.7       |4.6      |57       |

**Experiment 1**: Odds ratios for group comparisons. Odds ratios greater than 1 indicate increased odds of falling into a worse accuracy category compared to the kWh condition.

@@

As shown in @tbl-s1-reg, the reference class coefficients are positive for both the Percentage (Estimate = 1.3, 95% CI: 0.01 to 2.66, pd = 0.98) and USD (Estimate = 2.8, 95% CI: 1.52 to 4.04, pd = 1.00) conditions, relative to the kWh baseline. This indicates that, compared to the kWh condition, participants in both the Percentage and USD conditions were more likely to produce plans that fell into higher error categories. Moreover, the odds ratios (see Table 3) suggest that the USD condition led to a notably higher likelihood of large errors compared to the kWh baseline (OR = 15.7), while the Percentage condition also demonstrated increased odds (OR = 3.7) but was somewhat less detrimental to accuracy than USD. These results align with our descriptive findings and further clarify that framing the target reductions in absolute kWh units may facilitate significantly more accurate planning. Posterior predictive checks showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-ppd).

```{r}
#| label: fig-s1-ppd
#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal.  A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."

@@

## Experiment 1: Discussion

Experiment 1 examined how different numerical representations of energy reduction goals influenced participants’ planning accuracy. In line with our hypothesis that absolute units would yield better accuracy, the kWh condition supported significantly more precise energy reduction plans than did either the Percentage or USD conditions. Although the Percentage format was detrimental to accuracy relative to kWh, it was the USD condition that consistently produced the poorest outcomes, suggesting that monetary terms, while intuitive in everyday contexts, may not serve as effective reference classes for planning appliance-specific reductions in energy use.

Experiment 2 will extend these findings by examining whether additional variables, such as the difficulty of the reduction goal or the rounding of numerical values, further interact with reference class conditions, thereby providing a more comprehensive understanding of how to optimize energy information presentation for improved planning accuracy.

@@

```{r}
s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))
```

```{r}
#| eval: false
##| label: tbl-s2-agg
##| tbl-cap: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
